---
title: "Quality Control Alignments"
author: "Sergio E. Palma-Vera"
output: github_document
---

# Load libraries
```{r, message=F}
library(dplyr)
library(here)
library(stringr)
library(ggplot2)
library(reshape2)
library(tidyr)
```

# Load sample information
```{r}
sample_info <- read.csv(here("sample_info","sample_info_batch3","sample_info.csv"), header = T, stringsAsFactors = F) %>% mutate(name_short = str_remove(name, "-S1")) 

sample_info$Linie[sample_info$Linie == "HLB"] <- "DUHLB"

line_lev <- c("DUK","DUC","DU6","DU6P","DUHLB","FZTDU")

sample_info %>% head()
sample_info %>% dim()
```

# Load CollectWgsMetrics summary
```{r}
fls <- list.files(here("batch3/06_quality_control_alignments/output"), pattern = "/*_pt1_points.txt")

cvg_summary <- lapply(fls, function(fl){
  
#fl=fls[1]
  # Prepare sample name (taking into account droput-reseqd sample)
  cond <- fl == "I34772-L1_S19_L004.sorted.RG.dedup.bqsr.CollectWgsMetrics.default_pt1_points.txt" | fl == "I34772-L1_S19_L004.sorted.RG.dedup.bqsr.CollectWgsMetrics.Q0.M0_pt1_points.txt"
  if(cond){
    s <- "I34772_dropout_reseqd"
  }else{
    s <- str_split(fl,"-") %>% sapply(function(x) x[1])
  }

  # Load sample histogram information
  d <- read.table(file.path(here("batch3/06_quality_control_alignments/output"),fl), header = T, stringsAsFactors = F)
  
  # Extract mode of metrics (with or without bp and mapping quality filters)
  m <- str_split(fl,"_") %>% sapply(function(x) x[3]) %>% 
    str_remove("sorted.RG.dedup.bqsr.CollectWgsMetrics.") %>% 
    str_remove("L00[1-9].")
  
  # Add sample and mode information (default = mapping and bp quality of min 20)
  d <- mutate(d, sample = s, mode = m, mode = str_replace(mode, "default","Q20.M20"))
  
  # add mouse line information
  d <- dplyr::select(sample_info, Linie, name_short) %>% 
    right_join(d, by = c("name_short"="sample")) %>% 
    dplyr::rename(sample_name = name_short)
  
}) %>% 
  # bind rows
  bind_rows()

head(cvg_summary)
dim(cvg_summary) #182 31
```

# Add mouse line group to drop out sample
```{r}
cvg_summary <- cvg_summary %>% 
  mutate(Linie = ifelse(sample_name == "I34772_dropout_reseqd", "DUK", Linie))
  
```


# What's the mean coverage per sample?

Mean coverage per sample as boxplot/violinplot
```{r, eval = F}
png(here("batch3/06_quality_control_alignments/figures/boxplot_mean_coverage_per_sample.png"),res = 300, units = "px", height = 2000, width = 2000)

cvg_summary %>% 
  ggplot(data = ., aes(x = mode, y = MEAN_COVERAGE)) +
    geom_violin() +
    geom_jitter(width = 0.05, alpha = 0.5, aes(color = Linie )) +
    geom_boxplot(alpha = 0.2, width = 0.05, outlier.shape = NA) +
    xlab(NULL) +
    geom_hline(yintercept = 5, color = "red", linetype = "dotted")

dev.off()
```

Mean coverage per sample as barplot
```{r, eval = F}
png(here("batch3/06_quality_control_alignments/figures/barplot_mean_coverage_per_sample.png"),res = 300, units = "px", height = 2000, width = 3500)

cvg_summary %>% 
  ggplot(data = ., aes(x = sample_name, y = MEAN_COVERAGE, fill = Linie)) +
    xlab(NULL) +
    geom_bar(stat = "identity")+
    geom_hline(yintercept = 5, color = "red", linetype = "dotted") +
    facet_wrap(~mode, nrow = 2) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust  = 0.5))

dev.off()

```

Summarize (excluding original dropout sample)
```{r}
cvg_summary %>% 
  filter(sample_name != "I34772") %>% 
  group_by(mode) %>% 
  summarise(n(), min(MEAN_COVERAGE), mean(MEAN_COVERAGE), median(MEAN_COVERAGE),max(MEAN_COVERAGE),
            sum(MEAN_COVERAGE < 5)) %>% knitr::kable()
```

```{r}
cvg_summary %>% 
  filter(sample_name != "I34772") %>% 
  group_by(Linie,mode) %>% 
  summarise(n(), min(MEAN_COVERAGE), mean(MEAN_COVERAGE), median(MEAN_COVERAGE),max(MEAN_COVERAGE),
            sum(MEAN_COVERAGE < 5)) %>% knitr::kable()
```

```{r}
cvg_summary %>% filter(grepl("I34772", sample_name)) %>% 
  dplyr::select(Linie, sample_name, MEAN_COVERAGE)
```


* In average the mean cvg per sample was between 7-8x (median is similar)

* However there is a number of samples with less than 5x avg-cvg (14 if M0-Q0, and 19 if M20-Q20)

* The line with the most number of samples below avg-cvg 5x was DU6 (10 if M0-Q0 and 13 if M20-Q20).

* The drop-out sample has avg-cvg of ~0.2x, after resequencing mean cvg post-alignment was less than 4 (3,18-3.54).

* Samples in FZTDU and DUHLB were consistently above 5x

# Remove dropout sample
```{r}
cvg_summary <- cvg_summary %>% filter(sample_name != "I34772")
```


# Compare mean cvg raw reads, corrected reads and aligned reads

Load coverage information before alignment (raw fastq and fastp-corrected fastqs) and combine avg-cvg's
```{r}
avg_cvg <- list(
  aligned_Q0M0 = cvg_summary %>% 
    filter(mode == "Q0.M0") %>% 
    dplyr::select(sample_name,Linie, MEAN_COVERAGE) %>% 
    dplyr::rename(avg_cvg = MEAN_COVERAGE),
  aligned_Q20M20 = cvg_summary %>% 
    filter(mode == "Q20.M20") %>% 
    dplyr::select(sample_name,Linie, MEAN_COVERAGE) %>% 
    dplyr::rename(avg_cvg = MEAN_COVERAGE),  
  corrected = readRDS(here("batch3/03_quality_control_analysis/r_objects/qc_fastp_stats_avg_cvg.rds")),
  raw = readRDS(here("batch3/03_quality_control_analysis/r_objects/qc_raw_stats_avg_cvg.rds"))
  ) %>% 
  bind_rows(.id = "cvg_class") %>% 
  mutate(cvg_class = factor(cvg_class, levels = c("raw","corrected","aligned_Q0M0","aligned_Q20M20")))

head(avg_cvg)  
```

Remove dropout sample before sequencing
```{r}
avg_cvg <- avg_cvg %>% 
  filter(sample_name != "I34772_dropout") 
```


Calculate global cvg (all samples in cvg subset) mean 
```{r}
global_cvg <- avg_cvg %>% 
  group_by(cvg_class) %>% 
  summarise(mean = mean(avg_cvg))
global_cvg %>% knitr::kable()
```

Organize group levels
```{r}
avg_cvg$Linie <- factor(avg_cvg$Linie, levels = c("DUK","DUC","DU6","DU6P","DUHLB","FZTDU"))
```


Visualize
```{r, eval = F}
png(here("batch3/06_quality_control_alignments/figures/avg_cvg_by_class.png"),res = 300, units = "px", height = 2000, width = 3000)

ggplot(data = avg_cvg, aes(x = Linie, y = avg_cvg)) +
    geom_violin() +
    geom_jitter(width = 0.05, alpha = 0.5) +
    geom_boxplot(alpha = 0.2, width = 0.3, outlier.shape = NA) +
    facet_wrap(~cvg_class, nrow=1) +
    geom_hline(data = global_cvg, aes(yintercept = mean), 
               color = "darkgreen", linetype = "dotdash", size = 1) +
    geom_hline(yintercept = 5, color = "red", linetype = "dotted", size = 1) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    scale_y_continuous(breaks = seq(0,20,1)) +
    xlab(NULL)

dev.off()
```

* The mean coverage across all samples ranges from ~9x when data is raw tp 7.1x when reads had been aligned and filter by quality.

* Read-cleaning (correction) has a slight effect on global avg-cvg (9.02 vs 8.73)

* Using only high quality base pairs (M20Q20) reduces the aligned global avg-cvg slightly from 7.88x to 7.16x.

# Inspect the fraction of bases that attained at least X sequence coverage
```{r}
cols <- names(cvg_summary)[grep("^PCT_[[:digit:]]",names(cvg_summary))] # get cols with cvg
```


```{r, eval = F}
bks <- str_remove(cols, "PCT_") %>% str_remove("X")

png(here("batch3/06_quality_control_alignments/figures/min_cvg_genome_pct.png"),res = 300, units = "px", height = 2000, width = 2000)

cvg_summary %>% 
  dplyr::select(mode, Linie, sample_name, cols) %>% 
  melt(id.vars = c("mode", "Linie","sample_name")) %>% 
  mutate(value = as.numeric(value),
         cvg = str_remove(as.character(variable), "PCT_"),
         cvg = str_remove(cvg, "X"),
         cvg = as.numeric(cvg)) %>% 
  filter(cvg <= 20) %>% 
  ggplot(aes(x = cvg, y = value, color = sample_name, group = sample_name)) +
    geom_line() +
    facet_grid(Linie ~ mode) +
    theme(legend.position = "none") +
    geom_vline(xintercept = 5, color = "red", linetype = "dotdash") +
    scale_x_continuous(breaks = 1:20)

dev.off()
```

Faction of genome covered by at least 5x
```{r}
cvg_summary$Linie <- factor(cvg_summary$Linie, 
                            levels = c("DUK","DUC","DU6","DU6P","DUHLB","FZTDU"))

cvg_summary %>% 
  dplyr::select(mode, Linie, sample_name, cols) %>%
  melt(id.vars = c("mode", "Linie","sample_name")) %>%
  filter(variable == "PCT_5X") %>% 
  group_by(mode, Linie) %>% 
  summarise(n=n(), min=min(value), mean=mean(value), median=median(value), max=max(value)) %>% 
  knitr::kable(digits = 2)

cvg_summary %>% 
  dplyr::select(mode, Linie, sample_name, cols) %>%
  melt(id.vars = c("mode", "Linie","sample_name")) %>%
  filter(variable == "PCT_5X") %>% 
  group_by(mode) %>% 
  summarise(n=n(), min=min(value), mean=mean(value), median=median(value), max=max(value)) %>% 
  knitr::kable(digits = 2)
```

```{r, eval=F}
png(here("batch3/06_quality_control_alignments/figures/pct_territory_min_5x.png"),res = 300, units = "px", height = 2000, width = 3000)

cvg_summary %>% 
  dplyr::select(mode, Linie, sample_name, cols) %>%
  melt(id.vars = c("mode", "Linie","sample_name")) %>%
  filter(variable == "PCT_5X") %>% 
  ggplot(aes(Linie, value*100)) +
    geom_violin() +
    geom_jitter(width = 0.05, alpha = 0.5) +
    geom_boxplot(alpha = 0.2, width = 0.3, outlier.shape = NA) +
    facet_wrap(~mode) +
    ylab("Percentage Genome") +
    xlab(NULL) +
    ggtitle("Percentage Genome Covered at least 5x")

dev.off()
```





* The average proportion of genome covered at least 5x in any sample was approx 70%. 

* Half of the genomes had more than 70% 5x-cvg.

* A few genomes had very good cvg with > 90% territory with at least 5x

# How many reads were lost after mapping?
Prepare data
```{r}
pct_prop_paired <- read.table(here("batch3/06_quality_control_alignments/output/flagstat_pct_properly_paired_reads.tab"))

names(pct_prop_paired) <- c("sample_name","pct")

pct_prop_paired <- pct_prop_paired %>% left_join(sample_info, by = c("sample_name"="name_short"))

pct_prop_paired <- pct_prop_paired %>% dplyr::select(Linie,sample_name, pct)

head(pct_prop_paired)

```

Visualize
```{r, eval = F}
png(here("batch3/06_quality_control_alignments/figures/pct_properly_paired_reads.png"),
    res = 300, units = "px", height = 1000, width = 3500)

ggplot(pct_prop_paired, aes(x=sample_name, y = pct, fill = Linie)) +
  geom_bar(stat = "identity") +
  geom_hline(yintercept = 100, linetype = "dotted", color = "red") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  scale_y_continuous(breaks = seq(0,100,5)) +
  ggtitle("Percentage of properly paired reads") +
  xlab(NULL)
dev.off()
```

```{r}
pct_prop_paired %>% summarise(mean(pct > 90), min(pct))

pct_prop_paired %>% filter(pct == min(pct_prop_paired$pct))
```


* >96% of samples had a proportion of properly paired reads > 90%

* The lowest sample was I34741 (85.7%). Still an acceptable level.

* The drop-out sample had at least good mapping pct.


# What size are the insert (interval between mapped pairs)
Load data
```{r}
fls <- list.files(here("batch3/06_quality_control_alignments/output"), pattern = "CollectInsertSizeMetrics_pt1.txt")

insert_size_summary <- lapply(fls, function(fl){
  
#fl=fls[1]
  # Prepare sample name
  
    cond <- fl == "I34772-L1_S19_L004.sorted.RG.dedup.bqsr.CollectInsertSizeMetrics_pt1.txt"
  if(cond){
    s <- "I34772_dropout_reseqd"
  }else{
    s <- str_split(fl,"-") %>% sapply(function(x) x[1])
  }

  # Load sample histogram information
  d <- read.delim(file.path(here("batch3/06_quality_control_alignments/output"),fl), 
                  header = T, stringsAsFactors = F, comment.char = "#")
  
  # Add sample and mode information (default = mapping and bp quality of min 20)
  d <- mutate(d, sample = s)
  
  # add mouse line information
  d <- dplyr::select(sample_info, Linie, name_short) %>% 
    right_join(d, by = c("name_short"="sample")) %>% 
    dplyr::rename(sample_name = name_short)
  
}) %>% 
  # bind rows
  bind_rows()

head(insert_size_summary)
dim(insert_size_summary) #90 25
```

Visualize
```{r, eval = F}
png(here("batch3/06_quality_control_alignments/figures/insert_size_mean.png"),
    res = 300, units = "px", height = 1000, width = 3500)

ggplot(insert_size_summary, aes(x=sample_name, y = MEAN_INSERT_SIZE, fill = Linie)) +
  geom_bar(stat = "identity") +
  geom_errorbar(aes(ymin = MEAN_INSERT_SIZE-STANDARD_DEVIATION, 
                    ymax = MEAN_INSERT_SIZE+STANDARD_DEVIATION), width = 0.2, alpha = 0.5) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  xlab(NULL) +
  ggtitle("Mean Insert Size per sample (+/- sd)")

dev.off()
```

* Most samples had average insert size of ~400bp

* This is a reasonable interval consistent with the library preparation

# Conclusions

* In average the mean cvg per sample was between 7-8x (median is similar). However there is a number of samples with less than 5x avg-cvg (14 if M0-Q0, and 19 if M20-Q20). The line with the most number of samples below avg-cvg 5x was DU6 (10 if M0-Q0 and 13 if M20-Q20). The drop-out sample has avg-cvg of ~0.2x, after resequencing mean cvg post-alignment was less than 4 (3.18-3.54). Samples in FZTDU and DUHLB were consistently above 5x

* The mean coverage across all samples ranges from ~9x when data is raw tp 7.1x when reads had been aligned and filter by quality. Correction/cleaning has a slight effect on global avg-cvg (9.02 vs 8.73). Using only high quality base pairs (M20Q20) reduces the aligned global avg-cvg slightly from 7.88x to 7.16x.

* The average proportion of genome covered at least 5x in any sample was approx 70%. Half of the genomes had more than 70% 5x-cvg. A few genomes had very good cvg with > 90% territory with at least 5x. 

* 90% of corrected reads were properly in >95% of the samples. The lowest sample was I34741 (85.7%). Still an acceptable level. The drop-out sample had at least good mapping pct.

* Most samples had average insert size of ~400bp. This is a reasonable interval consistent with the library preparation



