---
title: "Count DUK/DUC alleles in FZTDU genotyped with GIGAMUGA"
output: 
  html_document:
    toc: 5
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, include = FALSE)

options(scipen=999)

library(tidyverse)
library(here)
library(vroom)
library(GenomicRanges)
```


```{r import_genotype_matrix}

FZTDU_Fwd_SNP_matrix <- here("selection_experiment/Henry/DNA_FZTDU_Gen_37_SNP_chip","FZTDU_Fwd_SNP-matrix.txt") %>% vroom()

```

```{r import_SNP_information}

Marker_Giga_MUGA <- here("selection_experiment/Henry/DNA_FZTDU_Gen_37_SNP_chip","Marker_Giga_MUGA.csv") %>% vroom()

Marker_Giga_MUGA %>% dplyr::select(IlmnID, Name, GenomeBuild, Chr, MapInfo, Source)

```


# Prelude

GIGAMUGA markers ("Marker_Giga_MUGA.csv") are based on build version 37, they were converted to coordinates in build version 38 (ENSEMBL: "assembly converter"). This is necessary for compatibility with WGS analysis

```{r prepare_vcf_like_file_for_conversion_37_to_38, eval = FALSE}

Marker_Giga_MUGA %>% dplyr::select(Chr, MapInfo) %>% mutate(id_37 = paste0("buid37:",Chr,"_", MapInfo), tmp1 = paste(rep(".",5), collapse = " ")) %>% 
  filter(!is.na(Chr)) %>% 
  write.table(
     here("selection_experiment/data","Marker_Giga_MUGA_37.vcf"), col.names = FALSE, row.names = FALSE, quote = FALSE
  )

```

```{r process_results_conversion_37_to_38}

Marker_Giga_MUGA_buid_37_to_38 <- here("selection_experiment/data", "Marker_Giga_MUGA_37_to_38_ensembl_assembly_converter.vcf") %>% 
  vroom(col_names = FALSE) %>% 
  # fix build37 information
  mutate(tmp = str_remove(X3,"buid37:")) %>% 
  separate(col = tmp, sep = "_", into = c("build37_chr", "build37_pos")) %>% 
  dplyr::rename(build38_chr = X1, build38_pos = X2) %>% 
  dplyr::select(starts_with("build")) 

```


```{r check_point, eval = FALSE}

inner_join(Marker_Giga_MUGA, 
           Marker_Giga_MUGA_buid_37_to_38 %>% mutate(build37_pos = as.double(build37_pos)), 
           by = c("Chr"="build37_chr", "MapInfo"="build37_pos"))

```


```{r include_alleles_and_frequencies}

DUK_AF <- here("batches123_08_LDD_SFS/output","cohort_biallelicSNPs_VQSR95_PASS_withmissingness.filtered.allrecords.DUK.frq") %>% 
  vroom(col_types = c("CHROM"="c")) %>% 
  inner_join(Marker_Giga_MUGA_buid_37_to_38, by = c("CHROM"="build38_chr", "POS"="build38_pos")) %>% 
  # prepare alllele and frequency columns
  separate(col = `{ALLELE:FREQ}`, sep = "\t", into = c("tmp1","tmp2")) %>% 
  separate(col = tmp1, sep = ":", into = c("REF","REF_frq")) %>% 
  separate(col = tmp2, sep = ":", into = c("ALT","ALT_frq")) %>% #38,646 x 10
  inner_join(Marker_Giga_MUGA %>% 
               dplyr::select(IlmnID, Name, Chr, MapInfo) %>% 
               mutate(MapInfo = as.character(MapInfo)), 
             c("build37_chr"="Chr", "build37_pos"="MapInfo")) %>% #39,972 x 12
  # Define allele state
  mutate(state = ifelse(REF_frq == 0, "fixed_alt", NA),
         state = ifelse(REF_frq == 1, "fixed_ref", state),
         state = ifelse(REF_frq > 0 & REF_frq < 1, "polymorphic", state)) %>% 
  # prepare genotype
  mutate(genotype = ifelse(state == "fixed_alt", paste0(ALT,"",ALT), NA),
         genotype = ifelse(state == "fixed_ref", paste0(REF,"",REF), genotype),
         genotype = ifelse(state == "polymorphic", paste0(REF,"",ALT), genotype))

DUC_AF <- here("batches123_08_LDD_SFS/output","cohort_biallelicSNPs_VQSR95_PASS_withmissingness.filtered.allrecords.DUC.frq") %>% 
  vroom(col_types = c("CHROM"="c")) %>% 
  inner_join(Marker_Giga_MUGA_buid_37_to_38, by = c("CHROM"="build38_chr", "POS"="build38_pos")) %>% 
  # prepare alllele and frequency columns
  separate(col = `{ALLELE:FREQ}`, sep = "\t", into = c("tmp1","tmp2")) %>% 
  separate(col = tmp1, sep = ":", into = c("REF","REF_frq")) %>% 
  separate(col = tmp2, sep = ":", into = c("ALT","ALT_frq")) %>% #38,646 x 10
  inner_join(Marker_Giga_MUGA %>% 
               dplyr::select(IlmnID, Name, Chr, MapInfo) %>% 
               mutate(MapInfo = as.character(MapInfo)), 
             c("build37_chr"="Chr", "build37_pos"="MapInfo")) %>% #39,972 x 12
  # Define allele state
  mutate(state = ifelse(REF_frq == 0, "fixed_alt", NA),
         state = ifelse(REF_frq == 1, "fixed_ref", state),
         state = ifelse(REF_frq > 0 & REF_frq < 1, "polymorphic", state)) %>% 
  # prepare genotype
  mutate(genotype = ifelse(state == "fixed_alt", paste0(ALT,"",ALT), NA),
         genotype = ifelse(state == "fixed_ref", paste0(REF,"",REF), genotype),
         genotype = ifelse(state == "polymorphic", paste0(REF,"",ALT), genotype))
  
```

Number of alleles intersecting WGS SNP call set
```{r, include=TRUE}

bind_rows(DUK_AF, DUC_AF) %>% dplyr::select(CHROM, POS) %>% unique() %>% nrow()

```

Animals can be female or male. The original data does not indicate what sex gets each integer (1 or 2)
```{r animal_info, include=TRUE}
animal_info <- here("selection_experiment/Henry/DNA_FZTDU_Gen_37_SNP_chip","DNA_FZTDU_Gen_37_SNP_chip.csv") %>% 
  vroom() %>% 
  mutate(animal_id = paste0("mmu_",`lfd. Nr.`)) %>% 
  dplyr::select(animal_id, everything())

animal_info %>% group_by(Sex) %>% summarise(n())
```


# Rules for alllele counts

Alleles in DUK or DUC can be:

- Fixed REF

- Fixed ALT

- Polymorphic


The genotype for an allele in an FZTDU animal can be 

- homozygous ALT (alt/alt)

- homozygous REF (ref/ref)

- Heterozygous (ref/alt or alt/ref)


Therefore, for marker x of FZTDU animal m, get allele counts:

- ref/ref vs fixed ref = 2

- ref/ref vs fixed alt = 0

- ref/ref vs polymorphic = 2

```{r, include=TRUE, echo=TRUE}
(c("T","T") %in% c("T","T")) %>% sum()
(c("T","T") %in% c("G","G")) %>% sum()
(c("T","T") %in% c("T","G")) %>% sum()
```

-alt/alt vs fixed ref = 0

-alt/alt vs fixed alt = 2

-alt/alt vs polymorphic = 2

```{r include=TRUE, echo=TRUE}
(c("G","G") %in% c("T","T")) %>% sum()
(c("G","G") %in% c("G","G")) %>% sum()
(c("G","G") %in% c("T","G")) %>% sum()
```


-ref/alt vs fixed ref = 1

-ref/alt vs fixed alt = 1

```{r include=TRUE, echo=TRUE}
(c("T","G") %in% c("T","T")) %>% sum()
(c("T","G") %in% c("G","G")) %>% sum()
```

-ref/alt vs polymorphic:

  - 1 if only one allele is seen in duk

  - 2 if both alleles are seen in duk

  - 0 if none of the alleles are seen

```{r include=TRUE, echo = TRUE}
(c("T","G") %in% c("T","C")) %>% sum()
(c("T","G") %in% c("T","G")) %>% sum()
(c("T","G") %in% c("C","A")) %>% sum()
```
  
- same for alt/ref genotypes

```{r, echo = TRUE, echo = TRUE}
(c("G","T") %in% c("T","T")) %>% sum()
(c("G","T") %in% c("G","G")) %>% sum()
(c("G","T") %in% c("T","C")) %>% sum()
(c("G","T") %in% c("T","G")) %>% sum()
(c("G","T") %in% c("C","A")) %>% sum()
```

```{r get_gene_counts_and_export, eval = FALSE}

# markers
markers <- bind_rows(DUK_AF, DUC_AF) %>% dplyr::select(Name) %>% unique() %>% .$Name #%>% .[1:5]

length(markers) #38646

names(markers) <- paste0("marker", 1:length(markers))

get_allele_counts <- function(line_AF){
  
  lapply(1:length(markers), function(i){ # loop over each marker in markers
    
    m <- markers[i]
      
    # print marker number 
    print(names(markers)[i])
  
    # get FZTDU genotypes for marker
    genotypes <- FZTDU_Fwd_SNP_matrix %>% filter(`SNP Name` == m) %>% dplyr::select(-`SNP Name`) %>% as.character() %>% str_split("") 
    
    # Get WGS allele state
    ref_line_pop_genotype <- line_AF %>% filter(Name == m) %>% .$genotype %>% str_split("") %>% unlist()
    
    # get allele counts in FZTDU
    allele_counts <- lapply(genotypes, function(gt) (gt %in% ref_line_pop_genotype) %>% sum()) %>% unlist()
    
    # tidy up
    allele_counts %>% as.data.frame() %>% data.table::transpose() %>% mutate(m = m) %>% dplyr::select(m, everything())
  
  }) %>% 
  bind_rows() %>% 
  magrittr::set_colnames(names(FZTDU_Fwd_SNP_matrix))
  
  
}


get_allele_counts(DUK_AF) %>% write.csv(here("selection_experiment/data/allele_counts_vs_DUK.csv"))


get_allele_counts(DUC_AF) %>% write.csv(here("selection_experiment/data/allele_counts_vs_DUC.csv"))

```

```{r import_counts_and_add_genomic_information_to_markers}
allele_counts_vs_DUK <- here("selection_experiment/data/allele_counts_vs_DUK.csv") %>% vroom() %>% dplyr::select(-"...1") %>% 
  inner_join(
    DUK_AF %>% dplyr::select(CHROM, POS, Name) %>% unique(), 
    by = c("SNP Name"="Name")
    ) %>% 
  dplyr::select(CHROM, POS, `SNP Name`, everything())

allele_counts_vs_DUC <- here("selection_experiment/data/allele_counts_vs_DUC.csv") %>% vroom() %>% dplyr::select(-"...1") %>% 
  inner_join(
    DUK_AF %>% dplyr::select(CHROM, POS, Name) %>% unique(), 
    by = c("SNP Name"="Name")
    ) %>% 
  dplyr::select(CHROM, POS, `SNP Name`, everything())
```

# Define chromosome sections and apply Prof Reinsch's Procedure

Chromosome section is defined as a genomic region where consecutive markers are no more than 1Mb apart.

Distance of 1Mb is set according to LD in FZTDU (r2 < 0.25). See fig 1b in preprint https://www.biorxiv.org/content/10.1101/2021.05.28.446207v1.full.pdf+html

```{r define_chr_sections}

allele_counts_vs_DUK_reduced <- allele_counts_vs_DUK %>% 
  mutate(seqnames = CHROM, start = POS, end = POS) %>% 
  dplyr::select(seqnames, start, end) %>% 
  makeGRangesFromDataFrame() %>% 
  reduce(min.gapwidth = 1e6) 


allele_counts_vs_DUC_reduced <- allele_counts_vs_DUC %>% 
  mutate(seqnames = CHROM, start = POS, end = POS) %>% 
  dplyr::select(seqnames, start, end) %>% 
  makeGRangesFromDataFrame() %>% 
  reduce(min.gapwidth = 1e6) 

identical(allele_counts_vs_DUK_reduced, allele_counts_vs_DUC_reduced) #TRUE

identical(allele_counts_vs_DUK %>% dplyr::select(`SNP Name`), 
          allele_counts_vs_DUC %>% dplyr::select(`SNP Name`)) # TRUE

chr_sections <- allele_counts_vs_DUC_reduced

df_chr_sections <- data.frame(chr = seqnames(chr_sections), start = start(chr_sections), end = end(chr_sections)) %>% as_tibble()

```

There are `r df_chr_sections %>% nrow()` in total.


```{r define_animal_ids}

anim_ids <- allele_counts_vs_DUK %>% dplyr::select(-CHROM, -POS, -`SNP Name`) %>% names() %>% as.character()

```

The sum of allele counts within each section was computed for each animal. Results were exported.

```{r compute_and_export_count_sums_per_chr_section_DUK, eval=FALSE}

cts=allele_counts_vs_DUK #arg

length(anim_ids) #456

section_sum_cts_DUK <- lapply(1:length(anim_ids), function(i){

  # aa=anim_ids[25] #loop
  aa <- anim_ids[i]
  print(i)
  
  cts_anim <- cts %>% dplyr::select(CHROM, POS, aa) 
  
  sum_cts <- lapply(1:nrow(df_chr_sections), function(i){
    
    #i=17
    
    section <- df_chr_sections[i,]
    
    #print(section)
    
    cts_section <- cts_anim %>% filter(CHROM == section$chr, POS >= section$start, POS <= section$end) %>% dplyr::select(aa) %>% unlist()
  
    cts_sum <- cts_section %>% sum()
  
    }) %>% 
    unlist() 
  
    res <- data.frame(x = sum_cts) %>% as_tibble()
    
    names(res) <- paste0("mmu_",aa)
    
    return(res)

  })

section_sum_cts_DUK %>% 
  purrr::reduce(bind_cols) %>% 
  bind_cols(
    df_chr_sections %>% dplyr::select(chr, start, end) 
  ) %>% 
  dplyr::select(chr, start, end, everything()) %>%
  write.csv(here("selection_experiment/data/section_sum_cts_DUK.csv"))

```

```{r compute_and_export_count_sums_per_chr_section_DUC, eval=FALSE}

cts=allele_counts_vs_DUC #arg

length(anim_ids) #456

section_sum_cts_DUC <- lapply(1:length(anim_ids), function(i){

  # aa=anim_ids[25] #loop
  aa <- anim_ids[i]
  print(i)
  
  cts_anim <- cts %>% dplyr::select(CHROM, POS, aa) 
  
  sum_cts <- lapply(1:nrow(df_chr_sections), function(i){
    
    #i=17
    
    section <- df_chr_sections[i,]
    
    #print(section)
    
    cts_section <- cts_anim %>% filter(CHROM == section$chr, POS >= section$start, POS <= section$end) %>% dplyr::select(aa) %>% unlist()
  
    cts_sum <- cts_section %>% sum()
  
    }) %>% 
    unlist() 
  
    res <- data.frame(x = sum_cts) %>% as_tibble()
    
    names(res) <- paste0("mmu_",aa)
    
    return(res)

  })

section_sum_cts_DUC %>% 
  purrr::reduce(bind_cols) %>% 
  bind_cols(
    df_chr_sections %>% dplyr::select(chr, start, end) 
  ) %>% 
  dplyr::select(chr, start, end, everything()) %>%
  write.csv(here("selection_experiment/data/section_sum_cts_DUC.csv"))

```

```{r import_section_sum_counts}

section_sum_cts_DUK <- here("selection_experiment/data/section_sum_cts_DUK.csv") %>% vroom() %>% dplyr::select(-`...1`)

section_sum_cts_DUC <- here("selection_experiment/data/section_sum_cts_DUC.csv") %>% vroom() %>% dplyr::select(-`...1`)

```

The mean of per-section sum was calculated among all animals

```{r compute_section_mean}

section_sum_cts_mean_DUK <- section_sum_cts_DUK %>% 
  dplyr::select(chr, start, end) %>% 
  mutate(
    MEAN = section_sum_cts_DUK %>% dplyr::select(starts_with("mmu_")) %>% apply(1, mean)
  )
  
section_sum_cts_mean_DUC <- section_sum_cts_DUC %>% 
  dplyr::select(chr, start, end) %>% 
  mutate(
    MEAN = section_sum_cts_DUC %>% dplyr::select(starts_with("mmu_")) %>% apply(1, mean)
  )
```

The standard deviation of per-section sum was calculated among all animals

```{r compute_section_sd}

section_sum_cts_sd_DUK <- section_sum_cts_DUK %>% 
  dplyr::select(chr, start, end) %>% 
  mutate(
    SD = section_sum_cts_DUK %>% dplyr::select(starts_with("mmu_")) %>% apply(1, sd)
  )
  
section_sum_cts_sd_DUC <- section_sum_cts_DUC %>% 
  dplyr::select(chr, start, end) %>% 
  mutate(
    SD = section_sum_cts_DUC %>% dplyr::select(starts_with("mmu_")) %>% apply(1, sd)
  )
```

Sections were transformed into z-scores by substracting the per-animal section sum from the among-animals section mean, then dividing by the among-animals section SD.

```{r calculate_z_scores}

anim_ids2 <- paste0("mmu_",anim_ids)

# DUK
z_scores_DUK <- lapply(anim_ids2, function(aa){
  
  res <- ( section_sum_cts_DUK[[aa]] - section_sum_cts_mean_DUK$MEAN ) / section_sum_cts_sd_DUK$SD
  
  names(res) <- NULL

  res_df <- data.frame(x = res)
  
  names(res_df) <- aa
  
  return(res_df)
  
  }) %>% purrr::reduce(bind_cols)

z_scores_DUK %>% dim()

# DUC
z_scores_DUC <- lapply(anim_ids2, function(aa){
  
  res <- ( section_sum_cts_DUC[[aa]] - section_sum_cts_mean_DUC$MEAN ) / section_sum_cts_sd_DUC$SD
  
  names(res) <- NULL

  res_df <- data.frame(x = res)
  
  names(res_df) <- aa
  
  return(res_df)
  
  }) %>% purrr::reduce(bind_cols)


z_scores_DUC %>% dim()

```

Sites with a standard deviation equal to zero were excluded as they are not informative
```{r}
z_scores_DUK_non_zero_SD <- z_scores_DUK[section_sum_cts_sd_DUK$SD > 0,]

z_scores_DUC_non_zero_SD <- z_scores_DUC[section_sum_cts_sd_DUC$SD > 0,]

dim(z_scores_DUK_non_zero_SD)

dim(z_scores_DUC_non_zero_SD)
```


The sum of all z-scores per animal was computed

```{r}
z_scores_DUK_sum <- z_scores_DUK_non_zero_SD %>% 
  apply(2, sum) %>% 
  as.data.frame() %>% 
  tibble::rownames_to_column() %>% 
  magrittr::set_colnames(c("animal_id","z_score")) %>% 
  arrange(desc(z_score)) %>% 
  # add sex information
  inner_join(animal_info, by = "animal_id") %>% 
  dplyr::select(animal_id, z_score, Sex) %>% 
  as_tibble()

z_scores_DUC_sum <- z_scores_DUC_non_zero_SD %>% 
  apply(2, sum) %>% 
  as.data.frame() %>% 
  tibble::rownames_to_column() %>% 
  magrittr::set_colnames(c("animal_id","z_score")) %>% 
  arrange(desc(z_score)) %>% 
  # add sex information
  inner_join(animal_info, by = "animal_id") %>% 
  dplyr::select(animal_id, z_score, Sex) %>% 
  as_tibble()
```

```{r export_animal_zscores, eval = FALSE}

z_scores_DUK_sum %>% write.csv(here("selection_experiment/data","rank_zscores_vs_DUK.csv"))

z_scores_DUC_sum %>% write.csv(here("selection_experiment/data","rank_zscores_vs_DUC.csv"))  
```

```{r export_by_sex}
z_scores_DUK_sum %>% filter(Sex == 1) %>% write.csv(here("selection_experiment/data","rank_zscores_vs_DUK_sex1.csv"))
z_scores_DUC_sum %>% filter(Sex == 1) %>% write.csv(here("selection_experiment/data","rank_zscores_vs_DUC_sex1.csv"))

z_scores_DUK_sum %>% filter(Sex == 2) %>% write.csv(here("selection_experiment/data","rank_zscores_vs_DUK_sex2.csv"))
z_scores_DUC_sum %>% filter(Sex == 2) %>% write.csv(here("selection_experiment/data","rank_zscores_vs_DUC_sex2.csv"))
```


Distribution of z-scores

```{r display_distribution, include=TRUE}

list(
  Ref_DUK = z_scores_DUK_sum,
  Ref_DUC = z_scores_DUC_sum
  ) %>% 
  bind_rows(.id = "reference_line") %>% 
  ggplot(aes(x = z_score)) +
    geom_histogram() +
    theme_bw() +
    xlab(NULL) +
    facet_wrap(~reference_line, nrow = 2, scales = "free") 
    
```

# Check for chromosome sections in animals with very low ranks. 

### Lowest rank animals for DUK and their aggregated z-scores.

```{r include=TRUE}

lowest_DUK <- z_scores_DUK_sum %>% tail(5)

lowest_DUK
```

### Distribution of z-scores for each chr-section
```{r include=TRUE}

z_scores_DUK_non_zero_SD %>% 
  dplyr::select(lowest_DUK$animal_id) %>% 
  reshape2::melt(variable.name = "animal_id", value.name = "z_score") %>% 
  as_tibble() %>% 
  ggplot(aes(x = z_score)) +
    geom_histogram() +
    theme_bw(base_size = 12) +
    facet_wrap(~animal_id, ncol = 1, strip.position = "right") +
    ggtitle("Distribution of chr sections z-scores")

```

### Those chr-sections have also low aggregated allele counts
```{r include=TRUE}

section_sum_cts_DUK %>% 
  dplyr::select(lowest_DUK$animal_id) %>% 
  reshape2::melt(variable.name = "animal_id", value.name = "section_cts_sum") %>% 
  as_tibble() %>% 
  ggplot(aes(x = section_cts_sum)) +
    geom_histogram() +
    theme_bw(base_size = 12) +
    facet_wrap(~animal_id, ncol = 1, strip.position = "right") +
    ggtitle("Distribution of chr sections sum of allele counts")

```

### Distribution of the standard deviations of the aggregated allele counts (unstandardized) over all chr sections
```{r include=TRUE}
section_sum_cts_sd_DUK %>% 
  ggplot(aes(x = SD)) +
    geom_histogram() +
    theme_bw(base_size = 12)
```

### Distribution among animals of most negative (z-score) chr-sections
```{r include=TRUE, warning=FALSE, fig.height=10}

z_scores_DUK_non_zero_SD %>% 
  mutate(min_among_animals = apply(., 1, min)) %>% 
  as_tibble() %>% 
  arrange(desc(min_among_animals)) %>% 
  #dplyr::select(min_among_animals) %>% 
  tail(5) %>% 
  mutate(chr_section_dummy_name = paste("chr_section",1:5)) %>% 
  reshape2::melt(variable.name = "animal_id", value.name = "z_score") %>% 
  as_tibble() %>% 
  ggplot(aes(x = z_score)) +
    geom_histogram(bins = 50) +
    theme_bw(base_size = 12) +
    facet_wrap(~chr_section_dummy_name, ncol = 1, strip.position = "right", scales = "free") +
    ggtitle(label = "Distribution of chr-sections among animals", subtitle = "Chr sections containing lowest z-scores values in at least one animal")

```

