---
title: "Analysis of WGS DU-mice data"
output:
  flexdashboard::flex_dashboard:
    orientation: columns
---

```{r results="asis"}
cat("
<style>
caption {
      color: black;
      font-weight: bold;
      font-size: 1.0em;
    }
</style>
")
```

```{r, setup, include=F}
options(scipen = 999)
library(flexdashboard)
library(dplyr)
library(stringr)
library(ggplot2)
library(reshape2)
library(tidyr)
library(here)
library(plotly)
library(DT)
library(UpSetR)
library(data.table)
library(vroom)
library(RColorBrewer)
library(knitr)
library(ggcorrplot)
library(tibble)
library(png)
library(grid)
library(gridExtra)
library(fastqcr)
```

```{r sample_info_and_genome_length, include=F}
sample_info <- read.csv(here("sample_info/sample_info_batch1_batch2.csv")) %>% 
  dplyr::select(Linie, sample_id) %>% 
  mutate(target_cvg = "30x") %>% 
  bind_rows(
    read.csv(here("sample_info/sample_info_batch3/sample_info.csv")) %>% 
      dplyr::select(Linie,name) %>% 
      dplyr::rename(sample_id = name) %>% 
      mutate(sample_id = str_remove(sample_id, "-S1"), target_cvg = "5x")
  ) %>% 
  mutate(Linie = ifelse(Linie == "HLB", "DUhLB",Linie))

#http://www.ensembl.org/Mus_musculus/Info/Annotation 
genome_length <- 2730871774 # genome length (Golden Path Length	) 
```

```{r function_import_fastqc_reports, include=F}
get_fastqc_data <- function(qc.dir){
  qc <- qc_aggregate(qc.dir)
  qc.stats <- qc_stats(qc)
  
  fields <- str_split(qc.stats$sample, "_")
  samp <- sapply(fields, function(x) x[1])
  R_pair <- sapply(fields, function(x) x[4])
  
  qc.stats$sample_name <- samp
  qc.stats$R_pair <- R_pair
  
  qc.stats$pct.dup <- as.numeric(qc.stats$pct.dup)
  qc.stats$pct.gc <- as.numeric(qc.stats$pct.gc)
  qc.stats$tot.seq <- as.numeric(qc.stats$tot.seq)
  qc.stats$seq.length <- as.numeric(qc.stats$seq.length)

  res <- list(qc=qc, qc.stats=qc.stats)
  return(res)
}

```

```{r batch1_prealignment_raw, include=F}
# import
fastqc_raw_60samp_9Lanes <- get_fastqc_data(file.path(here(),"batch1/01_quality_control/res_concat"))

# Add a column with only sample name to qc.stats
fastqc_raw_60samp_9Lanes$qc.stats <- fastqc_raw_60samp_9Lanes$qc.stats %>% 
  mutate(sample_name = str_sub(sample, 1, 6))

# Add column for read pair
fastqc_raw_60samp_9Lanes$qc.stats <- fastqc_raw_60samp_9Lanes$qc.stats %>% 
  mutate(R_pair = str_sub(sample, 8, 9))

# Calculate coverage and GB for each sample
NR_RL_GB_CVG_60samp_9Lanes_raw <- lapply(unique(fastqc_raw_60samp_9Lanes$qc.stats$sample_name),
#x="H07738"
                               function(x){
                                 # subset for sample
                                 dat <- fastqc_raw_60samp_9Lanes$qc.stats %>% 
                                   filter(sample_name == x)
                                 # prepare results read number and read length
                                 res <- data.frame(sample = x,
                                                   NR_R1 = dat$tot.seq[dat$R_pair == "R1"],
                                                   NR_R2 = dat$tot.seq[dat$R_pair == "R2"],
                                                   RL_R1 = dat$seq.length[dat$R_pair == "R1"],
                                                   RL_R2 = dat$seq.length[dat$R_pair == "R2"])
                                 # add GB, CVG and genome length
                                 res <- res %>% 
                                   mutate(GB = ((NR_R1*RL_R1) + (NR_R2*RL_R2))*10^-9,
                                          CVG=GB/( genome_length*10^-9 ),
                                          genome_length = genome_length)
                                 return(res)
                               }) %>% do.call(bind_rows,.)

# add sample info
NR_RL_GB_CVG_60samp_9Lanes_raw <- inner_join(NR_RL_GB_CVG_60samp_9Lanes_raw,
                                  sample_info,
                                  by = c("sample"="sample_id"))

```

```{r batch1_prealignment_clean, include=F}
setwd(here("batch1/03_quality_control_analysis/output_n_bp"))

NR_GB_CVG_60samp_9Lanes_trimm <- lapply(NR_RL_GB_CVG_60samp_9Lanes_raw$sample, 
#sample="H07738"       
       function(sample){
        ## Define files
        fls <- list.files()[grep(sample, list.files())]
        fls_NR <- fls[grep("NR", fls)]
        fls_GB <- fls[grep("GB", fls)]
        
        ## Load and aggregate files
        NR_R1 <- lapply(fls_NR[grep("R1",fls_NR)], function(x) read.delim(x, header = FALSE)) %>% 
          do.call(bind_rows,.) %>% 
          summarise(total_reads = sum(V1))
        NR_R2 <- lapply(fls_NR[grep("R2",fls_NR)], function(x) read.delim(x, header = FALSE)) %>% 
          do.call(bind_rows,.) %>% 
          summarise(total_reads = sum(V1))
        GB_R1 <- lapply(fls_GB[grep("R1",fls_NR)], function(x) read.delim(x, header = FALSE)) %>% 
          do.call(bind_rows,.) %>% 
          summarise(GB = sum(V1)*10^-9)
        GB_R2 <- lapply(fls_GB[grep("R2",fls_NR)], function(x) read.delim(x, header = FALSE)) %>% 
          do.call(bind_rows,.) %>% 
          summarise(GB = sum(V1)*10^-9)

        # summarize results for sample
        res <- data.frame(sample = sample, 
                          NR_R1=NR_R1$total_reads, 
                          NR_R2=NR_R2$total_reads,
                          GB=GB_R1+GB_R2)
        res <- res %>% mutate(CVG=GB/(genome_length*10^-9))

      }
      ) %>% do.call(bind_rows, .)

NR_GB_CVG_60samp_9Lanes_trimm <- inner_join(NR_GB_CVG_60samp_9Lanes_trimm,
                                  sample_info,
                                  by = c("sample"="sample_id"))
```

```{r batch1_postalignment, include=F}
path <- file.path(here(),"batch1/06_quality_control_alignments/output")

fls <- list.files(path)

dat_list <- lapply(fls, function(x){ 
  read.delim(file.path(path, x), stringsAsFactors = FALSE, dec = ",", skip = 6, nrow = 1)
  })

names(dat_list) <- substr(x = fls, start = 1, stop = 9)

CVG_post_align_60samp_9Lanes <- dat_list[grep("NoFilters.txt",fls)] %>% 
  bind_rows(.id = "sample_name") %>% 
  mutate(sample_name = substr(sample_name, 1, 6)) %>% 
  inner_join(sample_info, by = c("sample_name"="sample_id"))


CVG_post_align_qual_60samp_9Lanes <- dat_list[grep("WithFilters.txt",fls)] %>% 
  bind_rows(.id = "sample_name") %>% 
  mutate(sample_name = substr(sample_name, 1, 6)) %>% 
  inner_join(sample_info, by = c("sample_name"="sample_id"))
```

```{r batch2_prealignment_raw, include=F}
fastqc_raw_10samp_10Lanes <- get_fastqc_data(file.path(here(),"batch2/01_quality_control/output_concat"))

fastqc_raw_10samp_10Lanes$qc.stats <- fastqc_raw_10samp_10Lanes$qc.stats %>% 
  mutate(sample_name = str_sub(sample, 1, 6))

fastqc_raw_10samp_10Lanes$qc.stats <- fastqc_raw_10samp_10Lanes$qc.stats %>% 
  mutate(R_pair = str_sub(sample, 19, 20))

NR_RL_GB_CVG_10samp_10Lanes_raw <- lapply(unique(fastqc_raw_10samp_10Lanes$qc.stats$sample_name),
#x="H07738"
                               function(x){ 
                                 # subset for sample
                                 dat <- fastqc_raw_10samp_10Lanes$qc.stats %>% 
                                   filter(sample_name == x)
                                 # prepare results read number and read length
                                 res <- data.frame(sample = x,
                                                   NR_R1 = dat$tot.seq[dat$R_pair == "R1"],
                                                   NR_R2 = dat$tot.seq[dat$R_pair == "R2"],
                                                   RL_R1 = dat$seq.length[dat$R_pair == "R1"],
                                                   RL_R2 = dat$seq.length[dat$R_pair == "R2"])
                                 # add GB, CVG and genome length
                                 res <- res %>% 
                                   mutate(GB = ((NR_R1*RL_R1) + (NR_R2*RL_R2))*10^-9,
                                          CVG=GB/( genome_length*10^-9 ),
                                          genome_length = genome_length)
                                 return(res)
                               }) %>% do.call(bind_rows,.)

NR_RL_GB_CVG_10samp_10Lanes_raw <- inner_join(NR_RL_GB_CVG_10samp_10Lanes_raw,
                                  sample_info,
                                  by = c("sample"="sample_id"))
```

```{r batch2_prealignment_clean, include=F}
setwd(file.path(here(),"batch2/03_quality_control_analysis/GB_NR"))

NR_GB_CVG_10samp_1lane_trimm <- lapply(NR_RL_GB_CVG_10samp_10Lanes_raw$sample, 
#sample="H07739"       
       function(sample){  
        ## Define files reseq
        fls <- list.files()[grep(sample, list.files())]
        fls_NR <- fls[grep("NR", fls)]
        fls_GB <- fls[grep("GB", fls)]
        
        ## Load and aggregate files
        NR_R1 <- lapply(fls_NR[grep("R1",fls_NR)], function(x) read.delim(x, header = FALSE)) %>% 
          do.call(bind_rows,.) %>% 
          summarise(total_reads = sum(V1))
        NR_R2 <- lapply(fls_NR[grep("R2",fls_NR)], function(x) read.delim(x, header = FALSE)) %>% 
          do.call(bind_rows,.) %>% 
          summarise(total_reads = sum(V1))
        GB_R1 <- lapply(fls_GB[grep("R1",fls_NR)], function(x) read.delim(x, header = FALSE)) %>% 
          do.call(bind_rows,.) %>% 
          summarise(GB = sum(V1)*10^-9)
        GB_R2 <- lapply(fls_GB[grep("R2",fls_NR)], function(x) read.delim(x, header = FALSE)) %>% 
          do.call(bind_rows,.) %>% 
          summarise(GB = sum(V1)*10^-9)

        # summarize results for sample
        res <- data.frame(sample = sample, 
                          NR_R1=NR_R1$total_reads, 
                          NR_R2=NR_R2$total_reads,
                          GB=GB_R1+GB_R2)
        res <- res %>% mutate(CVG=GB/(genome_length*10^-9))

      }
      ) %>% do.call(bind_rows, .)

# 1+9=10 lanes 
NR_GB_CVG_10samp_10Lanes_trimm <- inner_join(NR_GB_CVG_10samp_1lane_trimm,
                                             NR_GB_CVG_60samp_9Lanes_trimm,
                                             by = "sample",
                                             suffix = c("_1lane", "_9Lanes")) %>% 
                                  select(sample, 
                                         NR_R1_9Lanes, NR_R2_9Lanes, NR_R1_1lane, NR_R2_1lane,
                                         GB_9Lanes, GB_1lane,
                                         CVG_9Lanes, CVG_1lane) %>% 
                                  mutate(CVG_10Lanes = CVG_9Lanes + CVG_1lane) %>% 
                                  inner_join(sample_info,by = c("sample"="sample_id"))

```

```{r batch2_postalignment, include=F}
path <- file.path(here(),"batch2/06_quality_control_alignments/output")

fls <- list.files(path)

dat_list <- lapply(fls, function(x){ 
  read.delim(file.path(path, x), stringsAsFactors = FALSE, dec = ",", skip = 6, nrow = 1)
  })

names(dat_list) <- substr(x = fls, start = 1, stop = 9)

CVG_post_align_10samp_10Lanes <- dat_list[grep("NoFilters.txt",fls)] %>% 
  bind_rows(.id = "sample_name") %>% 
  mutate(sample_name = substr(sample_name, 1, 6)) %>% 
  inner_join(sample_info, by = c("sample_name"="sample_id"))


CVG_post_align_qual_10samp_10Lanes <- dat_list[grep("WithFilters.txt",fls)] %>% 
  bind_rows(.id = "sample_name") %>% 
  mutate(sample_name = substr(sample_name, 1, 6)) %>% 
  inner_join(sample_info, by = c("sample_name"="sample_id"))  
  
```

```{r batch3_prealignment_data, include=F}
# objects previously created, just importing
# follow link to see how objects were created

qc_raw_stats_avg_cvg <- readRDS(here("batch3/03_quality_control_analysis/r_objects/qc_raw_stats_avg_cvg.rds"))
qc_fastp_stats_avg_cvg <- readRDS(here("batch3/03_quality_control_analysis/r_objects/qc_fastp_stats_avg_cvg.rds"))
```

```{r batch3_postalignment_data, include=F}
fls <- list.files(here("batch3/06_quality_control_alignments/output"), pattern = "/*_pt1_points.txt")

cvg_summary <- lapply(fls, function(fl){
  
#fl=fls[1]
  # Prepare sample name (taking into account droput-reseqd sample)
  cond <- fl == "I34772-L1_S19_L004.sorted.RG.dedup.bqsr.CollectWgsMetrics.default_pt1_points.txt" | fl == "I34772-L1_S19_L004.sorted.RG.dedup.bqsr.CollectWgsMetrics.Q0.M0_pt1_points.txt"
  if(cond){
    s <- "I34772_dropout_reseqd"
  }else{
    s <- str_split(fl,"-") %>% sapply(function(x) x[1])
  }

  # Load sample histogram information
  d <- read.table(file.path(here("batch3/06_quality_control_alignments/output"),fl), header = T, stringsAsFactors = F)
  
  # Extract mode of metrics (with or without bp and mapping quality filters)
  m <- str_split(fl,"_") %>% sapply(function(x) x[3]) %>% 
    str_remove("sorted.RG.dedup.bqsr.CollectWgsMetrics.") %>% 
    str_remove("L00[1-9].")
  
  # Add sample and mode information (default = mapping and bp quality of min 20)
  d <- mutate(d, sample = s, mode = m, mode = str_replace(mode, "default","Q20.M20"))
  
  # add mouse line information
  d <- sample_info %>% 
    right_join(d, by = c("sample_id"="sample")) %>% 
    dplyr::rename(sample_name = sample_id)
  
}) %>% 
  # bind rows
  bind_rows()

# Add mouse line group to drop out sample and prepare columns for integration
cvg_summary <- cvg_summary %>% 
  mutate(Linie = ifelse(sample_name == "I34772_dropout_reseqd", "DUK", Linie))  %>% 
  # fix droput/resequenced sample
  mutate(target_cvg = ifelse(is.na(target_cvg), "5x", target_cvg))
  
```


```{r collect_cvg_data, include=F}

boosted_samples <- NR_RL_GB_CVG_10samp_10Lanes_raw$sample

cvg_dat_prealignment <- bind_rows(
  
  # raw
  bind_rows(
    NR_RL_GB_CVG_60samp_9Lanes_raw %>% 
      filter(!(sample %in% boosted_samples)) %>% 
      dplyr::select(target_cvg, Linie, sample, CVG),
    NR_RL_GB_CVG_10samp_10Lanes_raw %>% 
      dplyr::select(target_cvg, Linie, sample, CVG),
    qc_raw_stats_avg_cvg %>% 
      mutate(target_cvg = "5x") %>% 
      dplyr::rename(sample = sample_name, CVG = avg_cvg) %>% 
      dplyr::select(target_cvg, Linie, sample, CVG)
  ) %>% 
    mutate(data_set = "cvg_raw"),
  
  #clean
  bind_rows(
    NR_GB_CVG_60samp_9Lanes_trimm %>% 
      filter(!(sample %in% boosted_samples)) %>% 
      dplyr::select(target_cvg, Linie, sample, CVG),
    NR_GB_CVG_10samp_10Lanes_trimm %>% 
      dplyr::select(target_cvg, Linie, sample, CVG_10Lanes) %>% 
      dplyr::rename(CVG = CVG_10Lanes),
    qc_fastp_stats_avg_cvg %>% 
      mutate(target_cvg = "5x") %>% 
      dplyr::rename(sample = sample_name, CVG = avg_cvg) %>% 
      dplyr::select(target_cvg, Linie, sample, CVG)
  ) %>% 
    mutate(data_set = "cvg_clean")
)


cvg_dat_postalignment <- bind_rows(
  
  # raw
  bind_rows(
    CVG_post_align_60samp_9Lanes %>% filter(!(sample_name %in% boosted_samples)),
    CVG_post_align_10samp_10Lanes,
    cvg_summary %>% 
      filter(mode == "Q0.M0") %>% 
      dplyr::select(names(CVG_post_align_10samp_10Lanes)) 
  ) %>% mutate(data_set = "m0q0"),
  
  # qual
  bind_rows(
    CVG_post_align_qual_60samp_9Lanes %>% filter(!(sample_name %in% boosted_samples)),
    CVG_post_align_qual_10samp_10Lanes,
    cvg_summary %>% 
      filter(mode == "Q20.M20") %>% 
      dplyr::select(names(CVG_post_align_qual_10samp_10Lanes))
  ) %>% mutate(data_set = "m20q20")
)

# combine cvg data before and after alignment
cvg_dat <- bind_rows(
  
  cvg_dat_prealignment %>% 
    dplyr::select(target_cvg, data_set, Linie, sample, CVG) %>% 
    # remove extra sample in 5x data set (the one discarded, aka dropout)
    filter(!(sample == "I34772_dropout")) %>% 
    # re-label sample that sample
    mutate(sample = ifelse(sample == "I34772_dropout_reseq", "I34772", sample)) %>% 
    # rename sample col
    dplyr::rename(sample_name = sample),
  
  cvg_dat_postalignment %>% 
    dplyr::select(target_cvg, data_set, Linie, sample_name, MEAN_COVERAGE) %>% 
    dplyr::rename(CVG = MEAN_COVERAGE) %>% 
    # remove extra sample in 5x data set (the one discarded, aka dropout)
    filter(!(sample_name == "I34772")) 
  ) %>% 
  # rename data set information
  mutate(
    data_set = ifelse(data_set == "cvg_raw", "raw_reads",data_set),
    data_set = ifelse(data_set == "cvg_clean", "clean_reads",data_set),
    data_set = ifelse(data_set == "m0q0", "aligned_m0q0",data_set),
    data_set = ifelse(data_set == "m20q20", "aligned_m20q20",data_set)
  ) %>% 
  # order levels for visualization
  mutate(
    Linie = ifelse(Linie == "DUHLB", "DUhLB", Linie),
    Linie = factor(Linie, levels = c("FZTDU","DUK","DUC","DU6","DU6P","DUhLB")),
    data_set = factor(data_set, levels = c("raw_reads","clean_reads","aligned_m0q0","aligned_m20q20"))
  )
```



Read Alignment
===============

Column {.tabset}
------------------

### Mean Genomewide Coverage
```{r visualize_cvg_data}

cvg_dat_means <- cvg_dat %>% 
  group_by(target_cvg, data_set) %>% summarise(mean = mean(CVG)) 

cvg_dat %>% 
  ggplot(aes(x = Linie, y = CVG, fill = target_cvg)) +
    geom_boxplot() +
    geom_hline(data = cvg_dat_means, aes(yintercept = mean, color = target_cvg), linetype = "dashed") +
    facet_wrap(~data_set, nrow = 1) +
    scale_y_continuous(breaks = seq(0,80,5)) +
    theme_bw() +
    xlab(NULL) +
    ylab("Mean Genome Coverage") +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)
    ) 

```

<<<<<<< HEAD
### Percentage of Genome Territory Covered
```{r pct_genome_cvg_visualization}
genome_cvg_pct_dat <- 
  # prepare data
  cvg_dat_postalignment %>% 
    # select columns
    dplyr::select(target_cvg, data_set, Linie, sample_name, ends_with("X")) %>% 
    # convert to long table
    reshape2::melt(
      id.vars = c("target_cvg","data_set", "Linie", "sample_name"),
      value.name = "pct_cvg",
      variable.name = "cvg_class"
    ) %>% 
    # fix x-axis and <-axis
    mutate(
      cvg_class = str_remove(cvg_class, "PCT_"),
      cvg_class = factor(cvg_class, paste0(c(1, seq(5,30,5), seq(40,100,10) ), "X") ),
      pct_cvg = pct_cvg*100
      ) %>% 
    #remove dropout sample (keep only re-sequenced one)
    filter(!(sample_name == "I34772"))

p <- genome_cvg_pct_dat %>% 
      ggplot(aes(x = cvg_class, y = pct_cvg)) +
        geom_boxplot() +
        geom_point(alpha = 0.3) +
        facet_grid(target_cvg ~ data_set) +
        theme_bw() +
        theme(
          legend.position = "none",
          axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)
        ) +
        ylab("Percentage Genome") +
        xlab(NULL) +
        scale_y_continuous(breaks = seq(0,100,10))

ggplotly(p)
```

### Per-sample metrics

### Mean CVG summary table
```{r}
cvg_dat %>% 
  group_by(target_cvg, data_set) %>% 
  summarise(min = min(CVG), mean = mean(CVG), median = median(CVG), max = max(CVG)) %>% 
  kable(digits = 2)
```

### Percentage of Genome covered summary
```{r}
genome_cvg_pct_dat %>% 
  group_by(target_cvg, data_set, cvg_class) %>% 
  summarise(min = min(pct_cvg), mean = mean(pct_cvg), median = median(pct_cvg), max = max(pct_cvg)) %>% 
  # prepare datatable output
  datatable(
    rownames = F, 
    options = list(
      columnDefs = list(list(className = 'dt-center', targets = "_all")),
      pageLength = 100
      )
    ) %>% 
  formatRound(columns = c("min","mean","median","max"), digits=2)

```




=======
>>>>>>> fff49821806fffc1986ba7871150e93d60977088




