---
title: "WGS DU-mice"
output:
  flexdashboard::flex_dashboard:
    orientation: columns
---


```{r results="asis"}
cat("
<style>
caption {
      color: black;
      font-weight: bold;
      font-size: 1.0em;
    }
</style>
")
```

```{r, setup, include=F}
options(scipen = 999)
```

Intro
====================================

* This is the analysis of the WGS data of DU mouse lines

* For every DU-line (DUK, DUC, DU6, DU6P, DUhLB, FZTDU) genomes of 25 animals were sequenced. Total of 150 genomes.

* Out the 25 genomes in one line, 10 were sequenced at ~25x and 15 at ~5x.

* This document begins with the evaluation the alignment metrics and concludes with the identification of genomic regions of distinct genetic differentiation for each selected line.

Bash_scripts
====================================

Column {.tabset}
-----------------

### Bash scripts

**Intro**

Text Text Text TextText TextText TextText TextText TextText TextText TextText TextText TextText Text


### Quality control 1

````
insert script

````


### Trimming

````
insert script

````


### cont....

